Separation of an upwelling current bounding the Juan de Fuca Eddy, off Vancouver Island

### 2022-04-25 PW13Paper.tex
FIX
Abstract

@@

Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal moxing. A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response (\fref{fig:HydraulicsCanyon}) during onshore flow (17:58--21:00).  Note the tide here is largely diurnal, so this onslope flow only occurs once a day.  Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \cite{deweycrawford88}.  The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.  This water is rapidly mixed with surrounding water such that its strong offshore spice values are attenuated.

>>>>

Comprehensive water column observations of temperature, salinity, and oxygen on the productive southern Vancouver Island shelf offer a unique characterization of a number of processes affecting exchange on the shelf.  A semi-permanent cyclonic ``eddy'' occupies the region in the lee of a bank where the shelf widens abruptly.  The observations indicate that the water in this wide part of the shelf has distinct water mass properties, and is clearly a mixture of offshore water and water in a coastal current.  The water in the ``eddy'' is either rapidly mixed (high $\kappa$), or has a long residence time (large $\tau$), such that $(\kappa \tau)^{1/2} \sim 50 -- 100\ \mathrm{m}$.  The offshore edge of this ``eddy'' water was very abrupt, with distinct water properties showing little lateral mixing between offshore and ``eddy'' water, with the front no more than 1-km wide. Previous work had hypothesized the role of a spur canyon incising the shelf in feeding water to the ``eddy''; little evidence of this was found during the observations, however, these were collected during a relaxation and reversal of upwelling winds, which may have slowed the flow of water into the eddy.  Tidal observations in the spur canyon also show very strong hydraulic flows in the cross-canyon direction accompanied by very high localized mixing rates in 50-m hydraulic jumps possibly exceeding $\kappa \sim 1 \ \mathrm{m^2s^{-1}}$.  Upstream of the observation there is evidence of an along-shelf current flowing equatorward.  This water is readily identified by partially mixed water properties, accumulated while on the shelf.  This water is clearly observed to all separate from the shelf to a depth of approximately 150 m, and is replaced offshore of the ``eddy'' by distinct offshore water.  This event can be seen in sea-surface temperatures in satellite images as tongue of water that is ejected offshore, probably with the relaxation of upwelling, with the flow heading directly offshore.  The cause of this offshore separation event are not determined, but likely include instability between the shelfbreak current and the California Undercurrent, and possible are catalyzed by separation from the abrupt underwater bank at the polward end of the study region.

@@

Much of the modified water in the Spur Canyon appears to come from the shelf to the west, but heavily modified by tidal moxing. A repeated tidal survey over the bank on the west side of the canyon shows a strong hydraulic response (\fref{fig:HydraulicsCanyon}) during onshore flow (17:58--21:00).  Note the tide here is largely diurnal, so this onslope flow only occurs once a day.  Given the stratification of $N\approx 6\times10^{-3} \ \mathrm{rad\,s^{-1}}$ and an overturning scale of 50 m, we might expect dissipation rates reaching $\epsilon \sim L^2N^{3} \approx 5\times10^{-4}\ \mathrm{m^2\,s^{-3}}$, which is more than three orders of magnitude higher than dissipation observed west of this location by \cite{deweycrawford88}.  In terms of a diapycnal diffusivity $\kappa = \gamma \epsilon / N^2 \approx 1\ \mathrm{m^2\,s^{-1}}$, assuming a mixing efficiency of $\gamma=0.2$ The effect on the properties in the canyon is for water to spill over from the offshore front into the canyon during the tide.  This water is rapidly mixed with surrounding water such that its strong offshore spice values are attenuated.
